### 2014-02-25 proposal/proposal.tex
Added LaTeX document for project proposal

\documentclass[11pt, oneside, fullpage, doublespace]{article}
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\usepackage{setspace}
\setstretch{1.5}

\title{CSCI-598a Project Proposal}
\author{Kevin Emery, Santiago Gonzalez, Brandon Rodriguez, Taylor Sallee}
\date{March 17, 2014}

\begin{document}
\maketitle

\section{Project Description}
A description

\subsection{Introduction}
Introduction

\subsection{Related Work}
Fun work

\section{Proposed Work}
\subsection{Automobile Detection}
Santiago
\subsection{Web Application}
Taylor
\subsection{Raspberry Pi}
Kevin
\subsection{Server Processing}
Brandon
\subsection{Central Basestation}
Everyone

\section{Summary}
A summary

\section{References}
References with \textbf{BibTeX}

\end{document}