### 2014-03-07 proposal/proposal.tex
Added content to related work

\documentclass[11pt, oneside, fullpage, doublespace]{article}
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\usepackage{setspace}
\setstretch{1.5}

\title{Wireless Automobile Detection, License Plate Processing, and Data Availability Network Proposal}
\author{Kevin Emery, Santiago Gonzalez, Brandon Rodriguez, Taylor Sallee\\ \emph{Undergraduates, EECS Department, Colorado School of Mines}}
\date{March 17, 2014}

\begin{document}
\maketitle

\begin{abstract}
300 words or less. Also, we need to put the EECS department in the title area as our professional affiliation.
\end{abstract}

\section{Project Description}
A description

\subsection{Introduction}
Introduction

\subsection{Related Work}
\cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using a magnetometer and wireless nodes based on the commercially available Arduino Fio microcontroller platform. This system, uses the ubiquitous IEEE 802.15.4 communications standard to communicate automobile detection data to a central base-station based on the small, commercially available Raspberry Pi Linux computer. The project described in this proposal will augment and extend this project while collaborating closely with Stillwell, to the point of a real world deployment at the Colorado School of Mines.

\section{Proposed Work}
\subsection{Automobile Detection}
Santiago
\subsection{Web Application}
Taylor
\subsection{Raspberry Pi}
Kevin
\subsection{Server Processing}
Brandon
\subsection{Central Basestation}
Everyone

\section{Summary}
A summary \cite{johnson2012}

\begin{thebibliography}{99}
\bibitem{stillwell2013} R. Stillwell, A. Wilson ``Magnetometer Parking Sensor'' \emph{EGGN 383 Final Project, Colorado School of Mines}. December 12, 2013
\bibitem{johnson2012} X. Johnson
\end{thebibliography}

\end{document}